---
title: "DSC 680 Project 1 R"
author: "Christine Hathaway"
date: "March 19, 2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include = TRUE}
#Set the working directory
setwd("C:/Users/Christine/Documents/Bellevue/DSC 680/Project 1")
```


```{r include = FALSE}
#Install packages needed for file
library(ggplot2)
library("plyr")
library(pastecs)
```

Import data from file

```{r echo = TRUE }
#Import file
flu <- read.csv(file = "C:/Users/Christine/Documents/Bellevue/DSC 680/Project 1/fluprint_export.csv", header = TRUE, na.strings = " ")
```

Display first five records of file

```{r echo = TRUE }
head(flu)
```

Run stat.desc() function of file

```{r echo = TRUE }
stat.desc(flu)
```


Run str() function of file

```{r echo = TRUE }
str(flu)
```

Count each variable

```{r echo = TRUE}
sapply(flu, function(x) length(unique(x)))

```

Unique values of variable "subset"

``` {r echo = TRUE}
table(flu$subset)

```

Unique values of variable "data"

``` {r echo = TRUE}
table(flu$data)

```

Create categorical variables for analysis
Variable are already factored, don't need this section

```{r echo = TRUE }

flu$Gender <- factor(flu$gender)

flu$Vaccine <- factor(flu$vaccine)

flu$Vaccine_Response <- factor(flu$vaccine_response)

flu$Influenza_Infection <- factor(flu$influenza_infection_history)

flu$Influenza_Hospitalization <- factor(flu$influenza_hospitalization)

flu$subtest <- factor(flu$subset)

head(flu)

```

Import data from file

```{r echo = TRUE }
#Import file
result <- read.csv(file = "C:/Users/Christine/Documents/Bellevue/DSC 680/Project 1/result_flu.csv", header = TRUE, na.strings = " ")
```

Display first five records of file

```{r echo = TRUE }
head(result)
```

Create categorical variables for analysis
Variable are already factored, don't need this section

```{r echo = TRUE }

result$Gender <- factor(result$gender)

result$Race <- factor(result$race)

result$Vaccine <- factor(result$vaccine)

result$Vaccine_Response <- factor(result$vaccine_response)

result$Influenza_Infection <- factor(result$influenza_infection_history)

result$Influenza_Hospitalization <- factor(result$influenza_hospitalization)

```

Scatterplot

``` {r echo = TRUE}
ggplot(flu, aes(x=data, y=vaccine_response)) + geom_point() + ggtitle("Data Value by Vaccine Response Type") +
  xlab("Data Values") + ylab("Vaccine Response")

```

``` {r echo = TRUE}
ggplot(flu, aes(x=subset, y=vaccine_response)) + geom_point()+ ggtitle("Test Types by Vaccine Response Type") +
  xlab("Test Type") + ylab("Vaccine Response")

```

Histograms

``` {r echo = TRUE}
ggplot(flu, aes(x=subset)) + geom_bar()+ ggtitle("Count of Various Test Types") +
  xlab("Test Types")

```

``` {r echo = TRUE}
ggplot(result, aes(x=Gender)) + geom_bar()+ ggtitle("Donors by Gender") +
  xlab("Gender")

```


``` {r echo = TRUE}
ggplot(result, aes(x=Race)) + geom_bar()+ ggtitle("Donors by Race") +
  xlab("Race")

```


``` {r echo = TRUE}
ggplot(result, aes(x=visit_age)) + geom_bar()+facet_wrap(~Vaccine_Response) + ggtitle("Donor Age by Vaccine Response") +
  xlab("Age")

```

``` {r echo = TRUE}
ggplot(result, aes(x=Gender)) + geom_bar()+facet_wrap(~Vaccine_Response)+ ggtitle("Gender by Vaccine Response Type") +
  xlab("Gender")

```


``` {r echo = TRUE}
ggplot(result, aes(x=visit_age)) + geom_bar() + ggtitle("Age Distribution") +
  xlab("Age")

```


Histograms

``` {r echo = TRUE}
ggplot(result, aes(x=Vaccine_Response)) + geom_bar() + ggtitle("Vaccine Response Distribution") +
  xlab("Response")

```



Create regression model for influenze infection based on vaccine type

```{r echo = TRUE }
model1 <- glm(Influenza_Infection~Vaccine+Gender+visit_age+Race, data = result, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model1)

```

Create regression model for vaccine response based on vaccine type

```{r echo = TRUE }
model2 <- glm(Vaccine_Response~Vaccine+Gender+visit_age+Race, data = result, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model2)

```


Create regression model for influenze infection based on donor factors

```{r echo = TRUE }
model12 <- glm(Influenza_Infection~Gender+visit_age+Race + bmi + statin_use, data = result, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model12)

```

Create regression model for vaccine response based on donor factors

```{r echo = TRUE }
model13 <- glm(Vaccine_Response~Gender+visit_age+Race + bmi + statin_use, data = result, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model13)

```


Regression model showing correlation between data values and vaccine response is significant

```{r echo = TRUE }
#Too big to run
model3 <- glm(vaccine_response~ data +gender+visit_age+race, data = flu, family = "binomial")

```


Output model results

```{r echo = TRUE }
summary(model3)

```

Names of tests with most frequent count of high vaccine response

``` {r echo = TRUE}
fluyes <- subset(flu, flu$vaccine_response == 1)
#head(fluyes)
rank <- sort(table(fluyes$name), decreasing = T)
head(rank)

```


Create subset of data based on CD8+ T cells

```{r echo = TRUE }
tcell8 <- subset(flu, name == 'CD8+ T cells')
head(tcell8)
str(tcell8)
```


Create regression model for vaccine response based on Test CD8+ T cells

```{r echo = TRUE }

model5 <- glm(vaccine_response~ data +gender+visit_age+race, data = tcell8, family = "binomial")

```


Output model results

```{r echo = TRUE }
summary(model5)

```

Create subset of data based on CD4+ T cells

```{r echo = TRUE }
tcell4 <- subset(flu, name == 'CD4+ T cells')
head(tcell4)
str(tcell4)
```


Create regression model for vaccine response based on Test CD4+ T cells 

```{r echo = TRUE }

model6 <- glm(vaccine_response~ data +gender+visit_age+race, data = tcell4, family = "binomial")

```


Output model results

```{r echo = TRUE }
summary(model6)

```

Create subset of data based on Tcells

```{r echo = TRUE }
tcell <- subset(flu, name == 'T cells')
head(tcell)
str(tcell)
```


Create regression model for vaccine response based on Test Tcells 

```{r echo = TRUE }

model7 <- glm(vaccine_response~ data +gender+visit_age+race, data = tcell, family = "binomial")

```


Output model results

```{r echo = TRUE }
summary(model7)

```

Create subset of data based on B Cells

```{r echo = TRUE }
bcells <- subset(flu, name == 'B cells')
head(bcells)
str(bcells)
```


Create regression model for vaccine response based on Test Bcells 

```{r echo = TRUE }

model8 <- glm(vaccine_response~ data +gender+visit_age+race, data = bcells, family = "binomial")

```


Output model results

```{r echo = TRUE }
summary(model8)

```

Create subset of data based on NK cells

```{r echo = TRUE }
nkcell <- subset(flu, name == 'NK cells')
head(nkcell)
str(nkcell)
```


Create regression model for vaccine response based on Test NK cells 

```{r echo = TRUE }

model9 <- glm(vaccine_response~ data +gender+visit_age+race, data = nkcell, family = "binomial")

```


Output model results

```{r echo = TRUE }
summary(model9)

```

