---
title: "DSC 680 Project 3 White Wine R"
author: "Christine Hathaway"
date: "May 10, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = TRUE}
#Set the working directory
setwd("C:/Users/Christine/Documents/Bellevue/DSC 680/Project 3")
```

```{r include = FALSE}
#Install packages needed for file
library(ggplot2)
library("plyr")
library(pastecs)
library(psych)
library("Hmisc")
library("caret")
library(dplyr)
library(gridExtra)
library(grid)
library(corrplot)
library(caTools)
```

Import data from file

```{r}
# Read data to ww dataframe
# row.names = 1 to avoid an index column creation upon dataset reading into a dataframe
ww <- read.csv('C:/Users/Christine/Documents/Bellevue/DSC 680/Project 3/winequality-white.csv', sep = ';')

```

Display first five records of file

```{r echo = TRUE }
head(ww)
```


Find dimensions of ww dataframe

```{r}
dim(ww)
```

List ww dataframe's column names, types and a subset of values

```{r}
str(ww)
```


Display summary statistics for each variable

```{r}
summary(ww)
```

Check how many missing values (NA) are in each column/variable, sum them up per column

```{r}
colSums(is.na(ww))
```


```{r}
# Draw a histogram for a given dataframe and variable
# Use deparse() and substitute() functions to decode column name from 
# a variable passed as an argument to the function, to be displayed
# on x axis (xlab())
draw_hist <- function(dataframe, variable)
{
  # Save histogram definition to the plot variable
  plot <- ggplot(data = dataframe, aes(x = variable)) + 
           geom_histogram(color = 'black', fill = '#099DD9') +
           xlab(deparse(substitute(variable)))
  return(plot)
}

# Build a matrix of small histograms with 3 columns
# using customly defined draw_hist() function
grid.arrange(draw_hist(ww, ww$fixed.acidity),
             draw_hist(ww, ww$volatile.acidity),
             draw_hist(ww, ww$citric.acid),
             draw_hist(ww, ww$residual.sugar),
             draw_hist(ww, ww$chlorides),
             draw_hist(ww, ww$free.sulfur.dioxide),
             draw_hist(ww, ww$total.sulfur.dioxide),
             draw_hist(ww, ww$density),
             draw_hist(ww, ww$pH),
             draw_hist(ww, ww$sulphates),
             draw_hist(ww, ww$alcohol),
             draw_hist(ww, ww$quality),
             ncol = 3)
            
```           

```{r}
# Plot a histogram of quality values
ggplot(data = ww, aes(x = quality)) +
  geom_histogram(color = 'black', fill = '#099DD9', binwidth = 1) +
  # Used to show 0-10 range, even if there are no values close to 0 or 10
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, 1)) +
  xlab('Quality of Red Wine') +
  ylab('Number of Red Wines')
  
```

````{r}
# Set boundaries for intervals
breaks <- c(0, 5, 7, 10)

# Bucket data points into intervals
ww$quality.category <- cut(ww$quality, breaks, include.lowest = TRUE, right = FALSE)

# Check intervals
summary(ww$quality.category)

# Add labels to intervals
labels <- c("Low", "Medium", "High")
ww$quality.category <- cut(ww$quality, breaks, include.lowest = TRUE, right = FALSE, labels=labels)

# Check if labels are applied properly
table(ww$quality.category)

```


```{r}
draw_boxplot <- function(dataframe, variable, ylab)
{
  plot <- ggplot(data = dataframe, aes(x = quality.category, y = variable, color = quality.category)) + 
    geom_boxplot() + 
    xlab('Quality') +
    #ylab(deparse(substitute(variable))) +
    ylab(ylab) +
    theme(legend.position = "none")
  return(plot)
}

# Build 4 boxplots summarizing distributions of 4 selected features
draw_univ_summary <- function()
{
  grid.arrange(draw_boxplot(ww, ww$volatile.acidity, expression(Volatile~Acidity~(g/dm^{3}))),
             draw_boxplot(ww, ww$citric.acid, expression(Citric~Acid~(g/dm^{3}))),
             draw_boxplot(ww, ww$sulphates, expression(Sulphates~(g/dm^{3}))),
             draw_boxplot(ww, ww$alcohol, 'Alcohol (% by Volume)'),
             ncol = 4,
             top = 'Features With Biggest Variability by Quality Category')
}

draw_univ_summary()


```


```{r}
# Create a new dataframe and calculate correlations 
# between ww variables
wwcor <- cor(ww[c(1:11, 12)])
# Draw a correlation matrix
corrplot(wwcor, method = 'square', order = "hclust", 
         tl.col = "black", tl.cex = 0.8, tl.offset = 1)

```


Regession models using binomial

```{r}
# create categorical variables

ww$category[ww$quality <= 5] <- 0
ww$category[ww$quality > 5] <- 1
ww$quality2 <- as.factor(ww$quality)

ww$category <- as.factor(ww$category)


head(ww)
```


Split data into Train Test sets

``` {r}

set.seed(3000)

spl = sample.split(ww$category, SplitRatio = 0.7)

wwtrain = subset(ww, spl==TRUE)
wwtest = subset(ww, spl==FALSE)

head(wwtrain)         
```

Create model

```{r}
model_glm <- glm(category ~ . - quality - quality2, data = wwtrain, family=binomial(link = "logit"))

```

Stepwise  model

```{r}
model_gl <- step(model_glm)

```

```{r}
head(fitted(model_gl))
```

```{r}
head(predict(model_gl))
```

```{r}
head(predict(model_gl, type = "response"))
```


Categorize wine

```{r}
trn_pred <- ifelse(predict(model_gl, type = "response") > 0.5,"Good Wine", "Bad Wine")
head(trn_pred)
```


Confusion matrix

```{r}
trn_tab <- table(predicted = trn_pred, actual = wwtrain$category)
trn_tab
```

Checking accuracy of the training set.

```{r}
sum(diag(trn_tab))/length(wwtrain$category)

```

Confusion matrix for the test data.

```{r}
# Making predictions on the test set.
tst_pred <- ifelse(predict(model_gl, newdata = wwtest, type = "response") > 0.5, "Good Wine", "Bad Wine")
tst_tab <- table(predicted = tst_pred, actual = wwtest$category)
tst_tab

```


Checking accuracy for the test data.

```{r}
sum(diag(tst_tab))/length(wwtest$category)
```



         
