---
title: "DSC 680 Project 2 R Code"
author: "Christine Hathaway"
date: "April 17, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = TRUE}
#Set the working directory
setwd("C:/Users/Christine/Documents/Bellevue/DSC 680/Project 2")
```


```{r include = FALSE}
#Install packages needed for file
library(ggplot2)
library("plyr")
library(pastecs)
library(psych)
library("Hmisc")
library("caret")
library(dplyr)
```

Import data from file

```{r echo = TRUE }
#Import file
df <- read.csv(file = "C:/Users/Christine/Documents/Bellevue/DSC 680/Project 2/parkinsons.csv", header = TRUE)
```


Display first five records of file

```{r echo = TRUE }
head(df)
```


Run str() function of file

```{r echo = TRUE }
str(df)
```

Count each variable

```{r echo = TRUE}
sapply(df, function(x) length(unique(x)))

```

Run desribe() function of file

```{r echo = TRUE }
describe(df)
```

Run stat.desc() function of file

```{r echo = TRUE }
stat.desc(df)
```

Create graphs of fundamental frequencies

```{r echo = TRUE}
ggplot(df, aes(x=MDVP.Fo.Hz.)) + geom_histogram() + ggtitle("Average vocal fundamental frequency")
ggplot(df, aes(x=MDVP.Fhi.Hz.)) + geom_histogram() + ggtitle("Maximum vocal fundamental frequency")
ggplot(df, aes(x=MDVP.Flo.Hz.)) + geom_histogram() + ggtitle("Minimum vocal fundamental frequency")

```


Create graphs of measures of variation in fundamental frequencies

```{r echo = TRUE}
ggplot(df, aes(x=MDVP.Jitter...)) + geom_histogram() + ggtitle("MDVP jitter in percentage")
ggplot(df, aes(x=MDVP.Jitter.Abs.)) + geom_histogram() + ggtitle("MDVP absolute jitter in ms")
ggplot(df, aes(x=MDVP.RAP)) + geom_histogram() + ggtitle("MDVP relative amplitude perturbation")
ggplot(df, aes(x=MDVP.PPQ )) + geom_histogram() + ggtitle("MDVP five-point period perturbation quotient")
ggplot(df, aes(x=Jitter.DDP )) + geom_histogram() + ggtitle("Average absolute difference of differences between jitter cycles")

```

Create graphs of measures of variation in amplitude

```{r echo = TRUE}
ggplot(df, aes(x=MDVP.Shimmer)) + geom_histogram() + ggtitle("MDVP local shimmer")
ggplot(df, aes(x=MDVP.Shimmer.dB.)) + geom_histogram() + ggtitle("MDVP local shimmer in dB")
ggplot(df, aes(x=Shimmer.APQ3)) + geom_histogram() + ggtitle("Three-point amplitude perturbation quotient")
ggplot(df, aes(x=Shimmer.APQ5)) + geom_histogram() + ggtitle("Five-point amplitude perturbation quotient")
ggplot(df, aes(x=MDVP.APQ)) + geom_histogram() + ggtitle("MDVP 11-point amplitude perturbation quotient")
ggplot(df, aes(x=Shimmer.DDA)) + geom_histogram() + ggtitle("Average absolute differences between the amplitudes of consecutive periods")

```

Create graphs of measures of ratio of noise to tonal components in the voice

```{r echo = TRUE}
ggplot(df, aes(x=NHR)) + geom_histogram() + ggtitle("Noise-to-harmonics ratio")
ggplot(df, aes(x=HNR)) + geom_histogram() + ggtitle("Harmonics-to-noise ratio")

```

Create graphs of nonlinear dynamical complexity measures

```{r echo = TRUE}
ggplot(df, aes(x=RPDE)) + geom_histogram() + ggtitle("Recurrence period density entropy measure")
ggplot(df, aes(x=D2 )) + geom_histogram() + ggtitle("Correlation dimension")

```

Create graphs of nonlinear dynamical complexity measures

```{r echo = TRUE}
ggplot(df, aes(x=spread1)) + geom_histogram() + ggtitle("Nonlinear measures of fundamental frequency variation")
ggplot(df, aes(x=spread2)) + geom_histogram() + ggtitle("Nonlinear measures of fundamental frequency variation")
ggplot(df, aes(x=PPE)) + geom_histogram() + ggtitle("Pitch period entropy")


```

Create graphs of Parkinson's status and signal fractal scaling exponent

```{r echo = TRUE}
ggplot(df, aes(x=status)) + geom_histogram() + ggtitle("Health status of the subject (one) - Parkinson's, (zero) - healthy")
ggplot(df, aes(x=DFA)) + geom_histogram() + ggtitle("Signal fractal scaling exponent")

```


Create Training and Testing Datasets

```{r echo = TRUE }
#create a list of random number ranging from 1 to number of rows from actual data and 70% of the data into training data  

data = sort(sample(nrow(df), nrow(df)*.7))

#creating training data set by selecting the output row values  
train<-df[data,]

#creating test data set by not selecting the output row values
test<-df[-data,]

```


Create regression model

```{r echo = TRUE }
model <- glm(status ~ NHR, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model)

```

Make predictions

```{r echo = TRUE }
#Make predictions
predict <- model %>% predict(test, type = "terms")

head(predict)

summary(predict)
```

Create regression model using all variables

```{r echo = TRUE }
model2 <- glm(status ~ MDVP.Fo.Hz.+MDVP.Flo.Hz.+MDVP.Jitter...+MDVP.Jitter.Abs.+MDVP.RAP + MDVP.PPQ + Jitter.DDP+ MDVP.Shimmer+ MDVP.Shimmer.dB.+Shimmer.APQ3 + Shimmer.APQ5 + MDVP.APQ +Shimmer.DDA +NHR+HNR+RPDE+ DFA+ spread1 + spread2 + D2 + PPE, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model2)

```

Make predictions

```{r echo = TRUE }
#Make predictions
predict <- model %>% predict(test, type = "terms")

head(predict)

summary(predict)
```



Create regression model using vocal fundamental frequency variables

```{r echo = TRUE }
model3 <- glm(status ~ MDVP.Fo.Hz.+MDVP.Flo.Hz.+MDVP.Jitter...+MDVP.Jitter.Abs.+MDVP.RAP+ MDVP.PPQ + Jitter.DDP+ MDVP.Shimmer+ MDVP.Shimmer.dB.+Shimmer.APQ3, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model3)

```


Create regression model using vocal fundamental frequency variables

```{r echo = TRUE }
model4 <- glm(status ~ Shimmer.DDA + HNR + NHR + DFA+RPDE+ spread1 + spread2 + D2 + PPE, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model4)

```

Create regression model using vocal fundamental frequency variables

```{r echo = TRUE }
model5 <- glm(status ~  Shimmer.APQ5 +MDVP.APQ, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(model5)

```
Based on XGBoost restuls in Python

```{r echo = TRUE }
modeltest <- glm(status ~   MDVP.Fo.Hz.+MDVP.Fhi.Hz.+Shimmer.APQ5+ spread1+ PPE, data = train, family = "binomial")

```

Output model results

```{r echo = TRUE }
summary(modeltest)

```

Make predictions

```{r echo = TRUE }
#Make predictions
predict <- modeltest %>% predict(test, type = "terms")

head(predict)

summary(predict)
```